### 2013-06-01 paper_draft.tex
Started to add text on FoldIt - fixed some minor words

\subsection*{Sharing genotypic information}
Currently users can upload their genotyping results from the companies \textit{23andMe}, \textit{deCODEme} and \textit{FamilyTreeDNA} via a web interface to the openSNP 
project. There is experimental support for uploading exomes in the VCF format \cite{Danecek01082011}, as \textit{23andMe} recently started exome sequencing for its customers. So far only the SNPs of the exome data sets are stored on openSNP, but the downloads include all variation found in the exome.
The uploaded data is published under the Creative Commons Zero license, 
which - in accordance with the Panton Principles \cite{10.1371/journal.pbio.1001195} - 
allows a complete re-use of the data without any constraints.
Between the start of openSNP on 09/27/2011 and 10/27/2012, 633 people have signed 
up with openSNP, and 270 genetic datasets have been made available. The openSNP 
database lists 215,546,685 genotypes which are distributed over 2,140,643 unique SNPs.
Figures \ref{Figure1_label} and \ref{Figure2_label} depict the increase in users and genotyping files since September 2011.

@@

PatientsLikeMe is a community for patients with life-changing illnesses to track and share the development of their illness with other patients with similar illnesses \cite{Wicks2010}.  This helps patients in understanding their diseases better - 72\% of surveyed participants found the site "moderately" or "very helpful", for example when it comes to starting a new medication (37\% found the site helpful), or when it comes to changing the medication (27\%). The data stored in PatientsLikeMe is open to the public and has been shown to be useful for research, for example in Multiple Sclerosis \cite{Bove2013}.

%COMPARE MORE. DO A TABLE. SHOULD THIS MAYBE BE PUT IN THE DISCUSSION INSTEAD OF RESULTS?

@@

\subsection*{GWAS and Open Data}
Although prices of exome or even full genome sequencing are dropping rapidly, GWAS are still considerably cheaper. However, GWAS can only detect correlations of SNPs with those traits and do not allow 
inference on the cause for any correlation. Furthermore, for a statistically sound analysis, GWAS need a large enough sample size. Nevertheless, GWAS are still frequently used and new associations are found \cite{10.1371.journal.pone.0031470,10.1371.journal.pone.0030309,10.1371.journal.pone.0029848}.

@@

\section*{Tables}
\begin{table}
\caption{Comparison of crowdsourced genetics platforms.}
\begin{tabular}{|l|l|l|l|l|l|}
\hline
Name & Provides annotation & Provides data & Open enrollment & API & License \\ 
\hline
SNPedia & \checkmark & \xmark & N/A & \checkmark & CC-NC-SA 3.0 \\ 
\hline
PGP & \checkmark & \checkmark & \xmark & \xmark & CC-BY 1.0, CC0 \\
\hline
PatientsLikeMe & \xmark & \checkmark & \checkmark & \xmark & CC-BY-SA 3.0 \\
\hline
openSNP & \checkmark & \checkmark & \checkmark & \checkmark & CC-BY 3.0 \\
\hline
\end{tabular}
\label{tab:platforms}
\end{table}
%\begin{table}
%\caption{Differences in terms of motivation to share genotypings with the public in survey-participants who already received a genotyping compared to participants who are not planning to getting genotyped. }  
%\begin{tabular}{|p{7cm}|p{2cm}|p{2cm}|p{2cm}|p{2cm}|}
%\hline
%& Mean \emph{Genotyped} & Mean \emph{Not Genotyped} & Mean Difference & Standard Error\\ 
%\hline
%\textbf{Motivation for sharing data in participants who are already genotyped} & & & & \\
%\hline 
%... curious & 3.82 & 2.66 & 1.159 & 0.193 \\ \hline % checked
%... want to help scientists & 4.64 & 4.18  & 0.465 & 0.128 \\ \hline % checked
%... for personal benefits & 3.77 & 3.32  & 0.448 & 0.183 \\ \hline % checked
%\textbf{Motivation for not sharing in participants who are not planning to get genotyped} & & & & \\ \hline
%... fear of discrimination & 3.09 & 4.15  & 1.06 & 0.195 \\ \hline % checked
%... breach of privacy & 3.01 & 3.68  & 0.666 & 0.211 \\ \hline % checked
%... fear of personalized advertising & 3.03 & 3.88  & 0.848 & 0.208 \\ \hline % checked
%... negative consequences for family members & 2.93 & 3.57  & 0.639 & 0.197 \\ \hline 
%\end{tabular}
%\label{tab:motivations1}
%\end{table}
%
%\begin{table}
%\caption{Differences in terms of motivations to share genotyping-data, comparison between participants who would share their genotyping data with participants who would not share their data.}
%\begin{tabular}{|p{7cm}|p{2cm}|p{2cm}|p{2cm}|p{2cm}|}
%\hline
%& Mean {Sharing} & Mean {Not Sharing} & Mean Difference & Standard Error \\ 
%\hline
%\textbf{Motivation for sharing genotypings in participants who would share} & & & & \\
%\hline
%... curiosity & 3.86 & 1.87 & 1.99 & 0.321 \\ \hline % checked
%... want to help science & 4.70 & 3.13 & 1.57 & 0.199 \\ \hline % checked
%... for personal benefits & 3.68 & 2.73 & 0.951 & 0.308 \\ \hline % checked
%\textbf{Motivation for sharing genotypings in participants who would not share} & & & & \\
%\hline
%... fear of discrimination & 3.21 & 4.73 & 1.52 & 0.322 \\ \hline % checked
%... fear of consequences for family members & 2.79 & 3.93 & 1.146 & 0.32 \\ \hline % checked
%... fear of personalized advertising & 3.17 & 4.29 & 1.112 & 0.357 \\ \hline % checked
%\end{tabular}
%\label{tab:motivations2}
%\end{table}
%\begin{table}[!ht]
%\caption{
%\bf{Table title}}
%\begin{tabular}{|c|c|c|}
%table information
%\end{tabular}
%\begin{flushleft}Table caption
%\end{flushleft}
%\label{tab:label}
% \end{table}

>>>>

\subsection*{Sharing genotypic information}
Currently users can upload their genotyping results from the companies \textit{23andMe}, \textit{deCODEme} and \textit{FamilyTreeDNA} via a web interface to the openSNP 
project. There is experimental support for uploading exomes in the VCF format \cite{Danecek01082011}, as \textit{23andMe} recently started exome sequencing for its customers. So far only the SNPs of the exome data sets are stored on openSNP, but the downloads include all variation found in the exome.
The uploaded data is published under the Creative Commons Zero license, 
which - in accordance with the Panton Principles \cite{10.1371/journal.pbio.1001195} - 
allows a complete re-use of the data without any constraints.
Between the launch of openSNP on 09/27/2011 and 10/27/2012, 633 people have signed 
up with openSNP, and 270 genetic datasets have been made available. As of 10/27/2012, the openSNP database lists 215,546,685 genotypes which are distributed over 2,140,643 unique SNPs.
Figures \ref{Figure1_label} and \ref{Figure2_label} depict the increase in users and genotyping files since September 2011.

@@

PatientsLikeMe is a community for patients with life-changing illnesses to track and share the development of their illness with other patients with similar illnesses \cite{Wicks2010}.  This helps patients in understanding their diseases better - 72\% of surveyed participants found the site "moderately" or "very helpful", for example when it comes to starting a new medication (37\% found the site helpful), or when it comes to changing the medication (27\%). Some subsets of data stored in PatientsLikeMe are open to the public and has been shown to be useful for research, for example in Multiple Sclerosis \cite{Bove2013}.

There are some projects that use gamification to let players crowdsource scientific data. For example, FoldIt is a puzzle game that lets players fold protein structures in order to achieve optimal structures. Players of Foldit have been able to identify protein structure and were even able to improve the activity of existing proteins structures \cite{Eiben2012}.
%COMPARE MORE. DO A TABLE. SHOULD THIS MAYBE BE PUT IN THE DISCUSSION INSTEAD OF RESULTS?

@@

\subsection*{GWAS and Open Data}
Although prices of exome or even full genome sequencing are dropping rapidly, GWAS are still considerably cheaper. However, GWAS can only detect correlations of SNPs with those traits and do not allow 
inference on the cause for any correlation. Furthermore, for a statistically sound analysis, GWAS need a large enough sample size. Nevertheless, GWAS are still frequently used and new associations are still being discovered \cite{10.1371.journal.pone.0031470,10.1371.journal.pone.0030309,10.1371.journal.pone.0029848}.

@@

\section*{Tables}
\begin{table}
\caption{Comparison of crowdsourced genetics platforms.}
\begin{tabular}{|l|l|l|l|l|l|}
\hline
Name & Provides annotation & Provides data & Open enrollment & API & License \\ 
\hline
SNPedia & \checkmark & \xmark & N/A & \checkmark & CC-NC-SA 3.0 \\ 
\hline
PGP & \checkmark & \checkmark & \xmark & \xmark & CC-BY 1.0, CC0 \\
\hline
PatientsLikeMe & \xmark & \checkmark & \checkmark & \xmark & Closed \& CC-BY-SA 3.0 \\
\hline
openSNP & \checkmark & \checkmark & \checkmark & \checkmark & CC-BY 3.0 \\
\hline
\end{tabular}
\label{tab:platforms}
\end{table}
%\begin{table}
%\caption{Differences in terms of motivation to share genotypings with the public in survey-participants who already received a genotyping compared to participants who are not planning to getting genotyped. }  
%\begin{tabular}{|p{7cm}|p{2cm}|p{2cm}|p{2cm}|p{2cm}|}
%\hline
%& Mean \emph{Genotyped} & Mean \emph{Not Genotyped} & Mean Difference & Standard Error\\ 
%\hline
%\textbf{Motivation for sharing data in participants who are already genotyped} & & & & \\
%\hline 
%... curious & 3.82 & 2.66 & 1.159 & 0.193 \\ \hline % checked
%... want to help scientists & 4.64 & 4.18  & 0.465 & 0.128 \\ \hline % checked
%... for personal benefits & 3.77 & 3.32  & 0.448 & 0.183 \\ \hline % checked
%\textbf{Motivation for not sharing in participants who are not planning to get genotyped} & & & & \\ \hline
%... fear of discrimination & 3.09 & 4.15  & 1.06 & 0.195 \\ \hline % checked
%... breach of privacy & 3.01 & 3.68  & 0.666 & 0.211 \\ \hline % checked
%... fear of personalized advertising & 3.03 & 3.88  & 0.848 & 0.208 \\ \hline % checked
%... negative consequences for family members & 2.93 & 3.57  & 0.639 & 0.197 \\ \hline 
%\end{tabular}
%\label{tab:motivations1}
%\end{table}
%
%\begin{table}
%\caption{Differences in terms of motivations to share genotyping-data, comparison between participants who would share their genotyping data with participants who would not share their data.}
%\begin{tabular}{|p{7cm}|p{2cm}|p{2cm}|p{2cm}|p{2cm}|}
%\hline
%& Mean {Sharing} & Mean {Not Sharing} & Mean Difference & Standard Error \\ 
%\hline
%\textbf{Motivation for sharing genotypings in participants who would share} & & & & \\
%\hline
%... curiosity & 3.86 & 1.87 & 1.99 & 0.321 \\ \hline % checked
%... want to help science & 4.70 & 3.13 & 1.57 & 0.199 \\ \hline % checked
%... for personal benefits & 3.68 & 2.73 & 0.951 & 0.308 \\ \hline % checked
%\textbf{Motivation for sharing genotypings in participants who would not share} & & & & \\
%\hline
%... fear of discrimination & 3.21 & 4.73 & 1.52 & 0.322 \\ \hline % checked
%... fear of consequences for family members & 2.79 & 3.93 & 1.146 & 0.32 \\ \hline % checked
%... fear of personalized advertising & 3.17 & 4.29 & 1.112 & 0.357 \\ \hline % checked
%\end{tabular}
%\label{tab:motivations2}
%\end{table}
%\begin{table}[!ht]
%\caption{
%\bf{Table title}}
%\begin{tabular}{|c|c|c|}
%table information
%\end{tabular}
%\begin{flushleft}Table caption
%\end{flushleft}
%\label{tab:label}
% \end{table}
